---
title: "ps_4"
author: "Niel Schrage"
date: "2/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
library(dplyr)
library(broom)
library(gganimate)
library(gifski)
library(gt)
library(stringr)

x<- read_csv("ps_4_elections-poll-nc09-3.csv",
  col_types = cols(
    .default = col_character(),
    turnout_scale = col_double(),
    turnout_score = col_double(),
    w_LV = col_double(),
    w_RV = col_double(),
    final_weight = col_double(),
    timestamp = col_datetime(format = "")))
```

## Question #1

```{r 1.1, echo= FALSE}
#select party, race, final weight then group by response and race.
a<- x %>% 
  filter(response == "Dem") %>% 
  nrow()
```
There were `r a` respondents who supported the Democratic candidate.

```{r 1.2, echo=FALSE}
a<- x %>% 
  filter(response %in% c("Und", "Rep")) %>% 
  group_by(response) %>% 
  count()
  
```
There were `r a[1,2] - a[2,2]` more respondents who favored the Republican candidate than who were Undecided.

```{r 1.3, echo= FALSE}
a<- x %>% 
  select(gender, gender_combined)
  
```
There are two gender variables (gender and gender_combined). There are X individuals for whom these variables have different values.

```{r 1.4}
#COME BACK
a<- x %>% 
  select(race_eth, file_race_black) %>% 
  filter(race_eth == "White")
  
```
There are X respondents listed as “White” under race_eth who are not listed as “White” under  file_race_black.

```{r 1.5}
a<- x %>% 
  select(response, timestamp) %>% 
  filter(response %in% c("Dem", "Rep")) %>% 
  arrange(timestamp) %>% 
  head()
  a
  
```
The first response of Rep came X minutes (rounded to the nearest minute) before the first response of Dem.

## Question #2
```{r 2}
b<- x %>% 
  select(response, race_eth, final_weight) %>% 
  filter(response %in% c("Rep", "Dem", "Und")) %>% 
  filter(race_eth %in% c("White", "Black", "Hispanic", "Asain", "Other")) %>% 
  group_by(response, race_eth)
```

## Question #3
```{r 3}

```

## Question #4
```{r 4}

```