---
title: "ps_4"
author: "Niel Schrage"
date: "2/25/2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load libraries
library(tidyverse)
library(dplyr)
library(broom)
library(gganimate)
library(ggplot2)
library(gifski)
library(gt)
library(stringr)
library(knitr)
library(ggthemes)

#read in data
x<- read_csv("ps_4_elections-poll-nc09-3.csv",
  col_types = cols(
    .default = col_character(),
    turnout_scale = col_double(),
    turnout_score = col_double(),
    w_LV = col_double(),
    w_RV = col_double(),
    final_weight = col_double(),
    timestamp = col_datetime(format = "")))
```

## Question #1

```{r 1.1, echo= FALSE}
#select party, race, final weight then group by response and race.
a<- x %>% 
  filter(response == "Dem") %>% 
  nrow()
```
There were `r a` respondents who supported the Democratic candidate.

```{r 1.2, echo=FALSE}
#filter by Und, Rep then group by response then just use basic math
b<- x %>% 
  filter(response %in% c("Und", "Rep")) %>% 
  group_by(response) %>% 
  count()
  
```
There were `r b[1,2] - b[2,2]` more respondents who favored the Republican candidate than who were Undecided.
```{r 1.3, echo= FALSE}
#select both variables then filter in the cases where gender does not equal gender_combined
c<- x %>% 
  select(gender, gender_combined) %>% 
  filter(gender != gender_combined) %>% 
  count()
  
```
There are two gender variables (gender and gender_combined). There are `r c[]` individuals for whom these variables have different values.

```{r 1.4, echo=FALSE}
#select both variables then filter in the cases where race does not equal race_file_black
d<- x %>% 
  select(race_eth, file_race_black) %>% 
  filter(race_eth == "White", file_race_black != "White") %>% 
  count()
  
```
There are `r d` respondents listed as “White” under race_eth who are not listed as “White” under  file_race_black.

```{r 1.5, echo=FALSE}
#selected party and timestamp, filtered by party, summarized, then spread then calcuated the difference. 
e<- x %>% 
  select(response, timestamp) %>% 
  filter(response %in% c("Dem", "Rep")) %>% 
  group_by(response) %>% 
  summarize(min = min(timestamp)) %>% 
  spread(response, min) %>% 
  mutate(diff = round(Rep - Dem, digits = 0))
  
```
The first response of Dem came `r e` minutes (rounded to the nearest minute) before the first response of Rep.

## Question #2
```{r 2 make_chart, echo= FALSE, results = "asis"}
x %>% 
  #selected by response, race, and weight, filtered data, releveled
  select(response, race_eth, final_weight) %>% 
  filter(response %in% c("Rep", "Dem", "Und", 3),
  race_eth != "[DO NOT READ] Don't know/Refused") %>% 
  mutate(race_eth = fct_relevel(race_eth, c("White", "Black", "Hispanic", "Asian","Other"))) %>% 
  group_by(response, race_eth) %>% 
  
  #calculated and mutated data to incorporate final weight
  summarize(total = sum(final_weight)) %>%
  
  #spread the data to get parties
  spread(key = response, value = total, fill = 0) %>%
  ungroup() %>% 
  mutate(all = Dem + Rep + Und + `3`) %>% 
  mutate(Dem = Dem/all) %>% 
  mutate(Rep = Rep/all) %>%
  mutate(Und = Und/all) %>% 
  select(-all, -`3`) %>% 
  
  #make na values 0 so that table isn't messed up
  na_if(0) %>% 
  
  #set up graph, labels, formatting by percent and missing values
  gt() %>% 
    tab_header(title = "Polling Results in North Carolina's 9th Congressional District") %>% 
    cols_label(race_eth = "Race", Dem = "DEM.",Rep = "REP.",Und = "UND.") %>%
    fmt_percent(columns = vars(Dem, Rep, Und),decimals = 0) %>% 
    fmt_missing(columns = 4, rows = 4, missing_text = "-") 

```

## Question #3
```{r 3, echo=FALSE}
#filter different education levels than reorder the data by education
x %>% 
  filter(educ %in% c("Bachelors' degree", "Grade school", "Graduate or Professional Degree", "High school", "Some college or trade school")) %>% 
  mutate(educ = fct_relevel(educ, c("Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))) %>% 
group_by(educ) %>% 

  #set up plot
  ggplot(aes(x = educ, y = final_weight)) + 
  coord_flip() + 
  #gets the violin shape
  geom_violin() + 
  
  #gets the dots shown
  geom_jitter(alpha = .5) + 
  theme_gray()+ 
  # style plot
  labs(title = "More Educated Matter Less in North Carolina 9th", 
       subtitle = "Poll gives more weight to people who are less likely to participate in polls", 
       caption = "Source: New York Times Upshot/Siena College 2018 live poll results") +
  xlab(element_blank()) +
  ylab("Weight Given to Respondent in Calculating Poll Results")

```

## Question #4
```{r 4, echo=FALSE}
#select variables -- question i want to answer: is there any relationship between political party and educational level
x %>% 
  filter(response %in% c("Rep", "Dem"), 
         educ %in% c("Bachelors' degree", "Grade school", "Graduate or Professional Degree", "High school", "Some college or trade school")) %>% 
  mutate(educ = fct_relevel(educ, c("Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))) %>%
  group_by(response, educ) %>% 
  summarize(total = sum(final_weight)) %>%
  
  #spread to find percents relative to total dems and reps in sample
  spread(key = response, value = total, fill = 0) %>%
  ungroup() %>% 
  mutate(all = Dem + Rep) %>% 
  mutate(Dem = Dem/all) %>% 
  mutate(Rep = Rep/all) %>%
  
  #basically reversing the spread so I can facet wrap. 
  gather(key = response, value = value, -educ) %>% 
  filter(response %in% c("Dem", "Rep")) %>% 
  
  #set up graph, dodge 2 makes it a stacked bar graph.
  ggplot(aes(fill = response, x = educ, y = value)) + 
    geom_col(position = "dodge2") +
    facet_wrap(~response) %>% 
    labs(title = "% of North Carolina District 9 Polled Voters Education by Party", 
         caption = "Theme: economist theme package, Data: Source: New York Times Upshot/Siena College") + 
    xlab(element_blank()) +
    ylab("Percentage") +
    theme_economist() +
    #changing x tick size
    theme(axis.text.x = element_text(size=7)) +
    scale_color_economist() +
    #recoloring so it matches Dem=blue, Rep=Red
    scale_fill_manual(values=c("blue", "red"))

```

## Collaboration
Worked with Mr. Gluck, Went to Mr. Flood's Office Hours. Sources: https://www.r-graph-gallery.com/48-grouped-barplot-with-ggplot2/, http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels?ref=driverlayer.com/web#change-the-appearance-of-the-axis-tick-mark-labels, 